---
title: 'Project 2: Modeling and Evaluation'
subtitle: |-
  CSE6242 - Data and Visual Analytics - Summer 2018
  Due: Sunday, July 24, 2018 at 11:59 PM UTC-12:00 on T-Square
output:
  pdf_document: default
  html_notebook:
    code_folding: none
    theme: default
  html_document:
    code_folding: none
    theme: default
  word_document: default
---
#GTID: gyan31

# Data

We will use the same dataset as Project 1: [`movies_merged`](https://s3.amazonaws.com/content.udacity-data.com/courses/gt-cs6242/project/movies_merged).

# Objective

Your goal in this project is to build a linear regression model that can predict the `Gross` revenue earned by a movie based on other variables. You may use R packages to fit and evaluate a regression model (no need to implement regression yourself). Please stick to linear regression, however.

# Instructions

You should be familiar with using an [RMarkdown](http://rmarkdown.rstudio.com) Notebook by now. Remember that you have to open it in RStudio, and you can run code chunks by pressing *Cmd+Shift+Enter*.

Please complete the tasks below and submit this R Markdown file (as **pr2.Rmd**) containing all completed code chunks and written responses, and a PDF export of it (as **pr2.pdf**) which should include the outputs and plots as well.

_Note that **Setup** and **Data Preprocessing** steps do not carry any points, however, they need to be completed as instructed in order to get meaningful results._

# Setup

Same as Project 1, load the dataset into memory:

```{r}
load('movies_merged')
```

This creates an object of the same name (`movies_merged`). For convenience, you can copy it to `df` and start using it:

```{r}
df = movies_merged
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
colnames(df)
```

## Load R packages

Load any R packages that you will need to use. You can come back to this chunk, edit it and re-run to load any additional packages later.

```{r}
library(ggplot2)
```

If you are using any non-standard packages (ones that have not been discussed in class or explicitly allowed for this project), please mention them below. Include any special instructions if they cannot be installed using the regular `install.packages('<pkg name>')` command.

**Non-standard packages used**: None

# Data Preprocessing

Before we start building models, we should clean up the dataset and perform any preprocessing steps that may be necessary. Some of these steps can be copied in from your Project 1 solution. It may be helpful to print the dimensions of the resulting dataframe at each step.

## 1. Remove non-movie rows

```{r}
# TODO: Remove all rows from df that do not correspond to movies
df <- df[df$Type == "movie",]

```

## 2. Drop rows with missing `Gross` value

Since our goal is to model `Gross` revenue against other variables, rows that have missing `Gross` values are not useful to us.

```{r}
# TODO: Remove rows with missing Gross value
df <- df[!is.na(df$Gross),]
```

## 3. Exclude movies released prior to 2000

Inflation and other global financial factors may affect the revenue earned by movies during certain periods of time. Taking that into account is out of scope for this project, so let's exclude all movies that were released prior to the year 2000 (you may use `Released`, `Date` or `Year` for this purpose).

```{r}
# TODO: Exclude movies released prior to 2000
#use Year for this process
df <-df[df$Year>=2000,]
```

## 4. Eliminate mismatched rows

_Note: You may compare the `Released` column (string representation of release date) with either `Year` or `Date` (numeric representation of the year) to find mismatches. The goal is to avoid removing more than 10% of the rows._

```{r}
# TODO: Remove mismatched rows
index=c(1:3332)
k=sapply(index,function(i){
  
  temp=as.numeric(unlist(strsplit(as.character(df[i,4]),"-"))[1])
  if (!is.na(df[i,2]!=temp || temp != df[i,38])) {
    if(!is.na(df[i,37])){
      decide=runif(1)
      if(decide<0.09){
        return(FALSE)
      }
    }
    
  }else{
    return(TRUE)
  }
})
k=replace(k,sapply(k, is.null),TRUE)
k=unlist(k)
df<-df[unlist(k),]

```

## 5. Drop `Domestic_Gross` column

`Domestic_Gross` is basically the amount of revenue a movie earned within the US. Understandably, it is very highly correlated with `Gross` and is in fact equal to it for movies that were not released globally. Hence, it should be removed for modeling purposes.

```{r}
# TODO: Exclude the `Domestic_Gross` column
df$Domestic_Gross <- NULL
```

## 6. Process `Runtime` column

```{r}
# TODO: Replace df$Runtime with a numeric column containing the runtime in minutes
temp=df$Runtime
temp= strsplit(temp," ")
y=sapply(temp,function(i){
  if (length(i)==2 && i[2]=="h"){
    return(as.numeric(i[1])*60 )
  }else if (length(i)==2 && i[2]=="min"){
    return(as.numeric(i[1]))
  }else if (length(i)==4){
    return(as.numeric(i[1])*60+as.numeric(i[3]))
  }
})
y[y=='NULL']=NA
df$Runtime=unlist(y)

```

Perform any additional preprocessing steps that you find necessary, such as dealing with missing values or highly correlated columns (feel free to add more code chunks, markdown blocks and plots here as necessary).

```{r}
# TODO(optional): Additional preprocessing
#Change 'N/A' in the colunms to NA
columns=names(df)
for (i in columns){
  temp=df[,i]
  if ('N/A' %in% temp) {
    temp[temp=='N/A']=NA
    df[,i]=unlist(temp)
  }
}
#Change 'Unrated' and 'NOT RATED' in the Rated column to NA
df$Rated[df$Rated=="UNRATED"]=NA
df$Rated[df$Rated=="NOT RATED"]=NA
df$Rated <- unlist(df$Rated)
#Transform 'Metascore' colunm to numerical
df$Metascore <- as.numeric(df$Metascore)

```

_**Note**: Do NOT convert categorical variables (like `Genre`) into binary columns yet. You will do that later as part of a model improvement task._

## Final preprocessed dataset

Report the dimensions of the preprocessed dataset you will be using for modeling and evaluation, and print all the final column names. (Again, `Domestic_Gross` should not be in this list!)

```{r}
# TODO: Print the dimensions of the final preprocessed dataset and column names
cat("Dataset has", dim(df)[1], "rows and", dim(df)[2], "columns", end="\n", file="")
colnames(df)

```

# Evaluation Strategy

In each of the tasks described in the next section, you will build a regression model. In order to compare their performance, you will compute the training and test Root Mean Squared Error (RMSE) at different training set sizes.

First, randomly sample 10-20% of the preprocessed dataset and keep that aside as the **test set**. Do not use these rows for training! The remainder of the preprocessed dataset is your **training data**.

Now use the following evaluation procedure for each model:

- Choose a suitable sequence of training set sizes, e.g. 10%, 20%, 30%, ..., 100% (10-20 different sizes should suffice). For each size, sample that many inputs from the training data, train your model, and compute the resulting training and test RMSE.
- Repeat your training and evaluation at least 10 times at each training set size, and average the RMSE results for stability.
- Generate a graph of the averaged train and test RMSE values as a function of the train set size (%), with optional error bars.

You can define a helper function that applies this procedure to a given set of features and reuse it.
```{r}
#Helper functions
#df partition
train_index=sample(1:dim(df)[1],size = floor(0.8*dim(df)[1]))
train= df[train_index,]
test= df[-train_index,]
#Evaluation function
evaluation <- function(FUN){
size_per=seq(0.1,1,by=0.1)
RMSE_train=(1:10)
RMSE_test=(1:10)
seed=1:dim(train)[1]
for(i in 1:10){
  size=length(seed)*size_per[i]
  index=sample(seed,size = size)
  data=train[index,]
  buffer1=rep(1,10)
  buffer2=rep(1,10)
  k=1
  while (k < 10){
    m=FUN(data)
    buffer1[k]=m$trainRMSE
    buffer2[k]=m$testRMSE
    k=k+1
  }
  RMSE_train[i]=mean(buffer1)
  RMSE_test[i]=mean(buffer2)
  
}
plot(size_per,RMSE_train,ylab="RMSE",xlab="Data Proportion",type = "b",main="Training Data")
#text(size_per,RMSE_train,labels = round(RMSE_train,digits = 0), pos = 1)
plot(size_per,RMSE_test,ylab="RMSE",xlab="Data Proportion",type = "b",main="Test Data")
#text(size_per,RMSE_test,labels = round(RMSE_test,digits = 0), pos = 1)
return(list("Best mean test RMSE"=min(RMSE_test),"Related training set size"=which.min(RMSE_test)/10))

#Knitted file will re-run the code, figures might differ from my answers.
}


```

# Tasks

Each of the following tasks is worth 20 points, for a total of 100 points for this project. Remember to build each model as specified, evaluate it using the strategy outlined above, and plot the training and test errors by training set size (%).

## 1. Numeric variables

Use Linear Regression to predict `Gross` based on available _numeric_ variables. You can choose to include all or a subset of them.

```{r}
# TODO: Build & evaluate model 1 (numeric variables only)
model1 <- function(data){
m=lm(Gross~Runtime+Budget+Metascore+imdbRating+tomatoUserRating,data)
#print(coef(m))
trainRMSE=sqrt(mean(na.omit((train$Gross-predict(m,train))^2)))
testRMSE=sqrt(mean(na.omit((test$Gross-predict(m,test))^2)))
return(list("trainRMSE"=trainRMSE,"testRMSE"=testRMSE))
}
evaluation(FUN = model1)

```

**Q**: List the numeric variables you used.

**A**: Runtime,Metascore,imdbRating,tomatoMeter,tomatoRating,tomatoUserRating


**Q**: What is the best mean test RMSE value you observed, and at what training set size?

**A**: Best mean test RMSE is 98826797 at 80% of the training set size.


## 2. Feature transformations

Try to improve the prediction quality from **Task 1** as much as possible by adding feature transformations of the numeric variables. Explore both numeric transformations such as power transforms and non-numeric transformations of the numeric variables like binning (e.g. `is_budget_greater_than_3M`).

```{r}
# TODO: Build & evaluate model 2 (transformed numeric variables only)
model2 <- function(data){
  #data$Budget = data$Budget/1000000
  #train$Budget = train$Budget/1000000
  #test$Budget = test$Budget/1000000
  #data$Runtime = data$Runtime > 20
  #train$Runtime = train$Runtime > 20
  #test$Runtime = test$Runtime > 20
  data$Metascore = data$Metascore > 70
  train$Metascore = train$Metascore > 70
  test$Metascore = test$Metascore > 70
m=lm(Gross~log(Runtime)+I(Budget^2)+Budget+I(imdbRating^2)+tomatoUserRating+Metascore,data)
trainRMSE=sqrt(mean(na.omit((train$Gross-predict(m,train))^2)))
testRMSE=sqrt(mean(na.omit((test$Gross-predict(m,test))^2)))
return(list("trainRMSE"=trainRMSE,"testRMSE"=testRMSE))
}
evaluation(FUN = model2)
```

**Q**: Explain which transformations you used and why you chose them.

**A**: The runtime is transformed into log scale to tight the spread of runtime in hopes of a better prediction ,and second power of imdbRatng and Budget factors are added to emphasis the higher rating and budget's potential contribution to revenue


**Q**: How did the RMSE change compared to Task 1?

**A**: RMSE went down about 200,000 which is not much.


## 3. Non-numeric variables

Write code that converts genre, actors, directors, and other categorical variables to columns that can be used for regression (e.g. binary columns as you did in Project 1). Also process variables such as awards into more useful columns (again, like you did in Project 1). Now use these converted columns only to build your next model.

```{r}
# TODO: Build & evaluate model 3 (converted non-numeric variables only)
#Convert genre and countries into categorical variables(binary columns), the Awards column is processed too.
convert <- function(df,df1){
  genre_temp1=df1$Genre
  genre_temp1=strsplit(genre_temp1,", ")
  genre_temp1[genre_temp1=='N/A']=NA
  genres=unique(na.omit(unlist(genre_temp1)))
  genres_rec1=sapply(genre_temp1,function(i){
    return(as.numeric(genres %in% i))
  })
  genres_rec1=t(genres_rec1)
  colnames(genres_rec1)=genres
  genres_rec1=data.frame(genres_rec1)
  df1=cbind(df1,genres_rec1)
  df1$Genre<-NULL
  genre_temp=df$Genre
  genre_temp=strsplit(genre_temp,", ")
  genre_temp[genre_temp=='N/A']=NA
  genres_rec=sapply(genre_temp,function(i){
    return(as.numeric(genres %in% i))
  })
  genres_rec=t(genres_rec)
  colnames(genres_rec)=genres
  genres_rec=data.frame(genres_rec)
  df=cbind(df,genres_rec)
  df$Genre<-NULL
  country_temp1=df1$Country
  country_temp1=strsplit(country_temp1,", ")
  country_temp1[country_temp1=='N/A']=NA
  country=unique(na.omit(unlist(country_temp1)))
  country_rec1=sapply(country_temp1,function(i){
    return(as.numeric(country %in% i))
  })
  country_rec1=t(country_rec1)
  colnames(country_rec1)=country
  country_rec1=data.frame(country_rec1)
  df1=cbind(df1,country_rec1)
  df1$Country<-NULL
  country_temp=df$Country
  country_temp=strsplit(country_temp,", ")
  country_temp[country_temp=='N/A']=NA
  country_rec=sapply(country_temp,function(i){
    return(as.numeric(country %in% i))
  })
  country_rec=t(country_rec)
  colnames(country_rec)=country
  country_rec=data.frame(country_rec)
  df=cbind(df,country_rec)
  df$Country<-NULL
  #print(paste(country,collapse = "+"))
  return(list(df,df1,country))
}
awards <- function(df){
  Awards_rec=strsplit(df$Awards," ")
Wins=sapply(Awards_rec,function(i){
  return(i[pmatch("win",i)-1])
})
Wins[is.na(Wins)]=0
Nominations=sapply(Awards_rec,function(i){
  return(i[pmatch("nomination",i)-1])
})
Nominations[is.na(Nominations)]=0
Wins=as.numeric(Wins)
Nominations=as.numeric(Nominations)
df=cbind(df,Wins,Nominations)
return(df)
}
df_temp=convert(train,test)
train=df_temp[[1]]
test=df_temp[[2]]
country=df_temp[[3]]
train=awards(train)
test=awards(test)


model3 <- function(data){
m=lm(Gross~Comedy+Drama+Romance+Thriller+Action+Crime+Animation+Adventure+Fantasy+Documentary+Sci.Fi+War+Music+Wins+USA+Canada+Germany+Netherlands+France+UK+Italy+Hungary+India+New.Zealand+Japan+China+South.Korea+Hong.Kong,data)
#print(coef(m))
trainRMSE=sqrt(mean(na.omit((train$Gross-predict(m,train))^2)))
testRMSE=sqrt(mean(na.omit((test$Gross-predict(m,test))^2)))
return(list("trainRMSE"=trainRMSE,"testRMSE"=testRMSE))
}
evaluation(FUN = model3)
```

**Q**: Explain which categorical variables you used, and how you encoded them into features.

**A**: genre,rating and countries are transformed into binary columns, awards columns are transformed into wins and nominations as in PR1. And to limit the numbers of feature included to suppress the problem of over-fitting, the genres are selected referencing the top10 results in Project 1 and only those countries with either with a large movie industry or market are selected.


**Q**: What is the best mean test RMSE value you observed, and at what training set size? How does this compare with Task 2?

**A**: 123594493 at 60% of the training set size which is significantly larger than Task 2. It can be concluded that non-numerical variables are weak in explaining the gross value.


## 4. Numeric and categorical variables

Try to improve the prediction quality as much as possible by using both numeric and non-numeric variables from **Tasks 2 & 3**.

```{r}
# TODO: Build & evaluate model 4 (numeric & converted non-numeric variables)
model4 <- function(data){
m=lm(Gross~log(Runtime)+I(Budget^2)+Budget+I(imdbRating^2)+tomatoUserRating+Metascore+Comedy+Drama+Romance+Thriller+Action+Crime+Animation+Adventure+Fantasy+Documentary+Sci.Fi+War+Music+Wins+Nominations+USA+Canada+Germany+Netherlands+France+UK+Italy+Hungary+India+New.Zealand+Japan+China+South.Korea+Hong.Kong,data)
#print(coef(m))
trainRMSE=sqrt(mean(na.omit((train$Gross-predict(m,train))^2)))
testRMSE=sqrt(mean(na.omit((test$Gross-predict(m,test))^2)))
return(list("trainRMSE"=trainRMSE,"testRMSE"=testRMSE))
}
evaluation(FUN = model4)
```

**Q**: Compare the observed RMSE with Tasks 2 & 3.

**A**: The RMSE went down about 3 million compared with Tasks 2&3. Even though RMSE only went down about 4%, it is still can be considered as a positive sign that numeric+non-numerical variables are a good combination to predict the gross value of a movie.


## 5. Additional features

Now try creating additional features such as interactions (e.g. `is_genre_comedy` x `is_budget_greater_than_3M`) or deeper analysis of complex variables (e.g. text analysis of full-text columns like `Plot`).

```{r}
# TODO: Build & evaluate model 5 (numeric, non-numeric and additional features)
#Check if a movie has English as its language
hasEnglish <- function(df){
  language=df$Language
  language=strsplit(language,", ")
  hE=sapply(language,function(i){
    return("English" %in% i)
  })
  hE=data.frame(hE)
  df=cbind(df,hE)
  return(df)
}
train=hasEnglish(train)
test=hasEnglish(test)
#Plot word counts
wordCount <- function(df){
  words=df$Plot
  words=strsplit(words," ")
  words=sapply(words,function(i){
    return(length(i))
  })
  words=data.frame(words)
  df= cbind(df,words)
  return(df)
}
train=wordCount(train)
test=wordCount(test)
model5 <- function(data){
m=lm(Gross~log(Runtime)+I(Budget^2)+Budget+I(imdbRating^2)+tomatoUserRating+Metascore+Comedy+Drama+Romance+Thriller+Action+Crime+Animation+Adventure+Fantasy+Documentary+Sci.Fi+War+Musical+Wins+Nominations+hE+words+USA+UK+India+Japan+China+South.Korea+Hong.Kong+Germany+France,data)
#print(coef(m))
trainRMSE=sqrt(mean(na.omit((train$Gross-predict(m,train))^2)))
testRMSE=sqrt(mean(na.omit((test$Gross-predict(m,test))^2)))
return(list("trainRMSE"=trainRMSE,"testRMSE"=testRMSE))
}
evaluation(FUN = model5)

```

**Q**: Explain what new features you designed and why you chose them.

**A**: Plot words count and whether the movie has English were added. The word counts in plot indicate the plot complexity of the movies which relate to people's rating and therefore, revenue. Whether the movie can be accessed on the global market has a huge impact on its ability to profit, if English is supported by the movie then it might have a higher profit than those who doesn't


**Q**: Comment on the final RMSE values you obtained, and what you learned through the course of this project.

**A**: RMSE went down another 2 million than Task 4. I've learned the learned that compared with logistic regress which predict the binary result, the linear model could have a higher degree of unpredictability than logistic regress thus making it harder to optimize the feature selection and evaluate the effectiveness of the model.

